Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses

### 2024-12-10 manuscript_agu.tex
Plain language summary

\section*{Plain Language Summary}
Enter your Plain Language Summary here or delete this section.
Here are instructions on writing a Plain Language Summary:
https://www.agu.org/Share-and-Advocate/Share/Community/Plain-language-summary

@@

Compared to lidar observations, the daily cloud cover tends to be about 1 okta lower in ICON and 2 oktas lower in the reanalyses. Conditions of weak stability are associated with some of the greatest biases, especially in ERA5. The models also underestimate the cloud cover very strongly in cyclonic conditions, which are very cloudy in the observations (8 oktas), but much less so in the models.

>>>>

\section*{Plain Language Summary}

Global storm-resolving models are climate models with km-scale horizontal resolution, which are currently in development. Reanalyses are the best estimates of past meteorological conditions based on an underlying global model and observations. We evaluated clouds and thermodynamic profiles over the Southern Ocean in one such model, as well as two reanalyses, based on 2400 days of ship and station observations. Thanks to the high resolution, the model relies entirely on explicit simulation of clouds, instead of subgrid-scale parametrizations. For the evaluation, we used ceilometer and radiosonde observations and a lidar simulator, which enables a fair comparison with the model and reanalyses. We subsetted our results by cyclonic activity and stability. We found that the model and reanalyses underestimate a lidar-derived cloud fraction, and the reanalyses do so more strongly. Fog or near-surface clouds are especially underestimated in the reanalyses. However, the model and one of the reanalyses also tend to overestimate the peak of cloud occurrence at 500 m above the ground, and it tends to be higher. This is linked to thermodynamic profiles, which show a higher lifting condensation level. Southern Ocean biases are still an important problem in the model, but an improvement over the reanalyses is notable.

@@

Compared to lidar observations, the daily cloud cover tends to be about 1 okta lower in ICON and 2 oktas lower in the reanalyses. Conditions of weak stability are associated with some of the greatest biases, especially in ERA5. The models also underestimate the cloud cover very strongly in cyclonic conditions, which are very cloudy in the observations (8 oktas), but much less so in the models. Similarly, \citeA{mcerlich2023} found a 40\% underestimation of cloud liquid water in cyclones over the SO in ERA5, despite total column water vapor simulated much more accurately (5\% underestimation).
